---
title: "CUNY DATA607_Wk9_API"
author: "Zachary Herold"
date: "October 28, 2018"
output: html_document
---

```{r message = FALSE, warning = FALSE}
library(httr)
library(jsonlite)
library(stringr)
library(pander)
```

## NYT API Access Denied

In Week 9's assignment, the task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.

I first checked NYTimes public specs on Github to find the API URL format. 

https://github.com/NYTimes/public_api_specs/commit/fca6f6c9def8eede59726f3b06a2734f07e689ee

My inital interest was in the Top Stories category. 

I set up a concatenated URL from the host, path, query and api-key for the API call. 

Here, the API-key is concatenated into the query directly, though I would rather use Set environment function (Sys.setenv) to hide secrets.

```{r}
host <- "http://api.nytimes.com"
path <- "/svc/topstories/v2.json"
query <- "?order=by-date"
api_key <- "&api-key=af67b77762da45ccbdcded839b22ce8a"

nyt.url <- str_c(host,path, query, api_key, sep = "", collapse = NULL)

nyt.url
```

The GET method in Wickham's httr package makes a request to NYT's server, only to be rejected.


```{r}
r <- GET(nyt.url, verbose())
## <- HTTP/1.1 403 Forbidden
```

HTTP/1.1 403 Forbidden
Access-Control-Allow-Credentials: false 


```{r}
http_status(r)
```

The Top Stories seem off limits. The New York Times website is off-limits in China, blocked by the Great Firewall. I suppose such a news feed too may be suppressed.


## Movie Requests and Bestseller Lists A-Ok

When we set the path to movie reviews, API access is suddenly permitted. 

```{r}
path <- "/svc/movies/v2/reviews/dvd-picks.json"
nyt.url <- str_c(host, path, query, api_key, sep = "", collapse = NULL)

r <- GET(nyt.url, verbose())
```

Booklists are apparently also not subject to the Great Firewall. 

```{r}
path <- "/svc/books/v2/lists/overview.json"
query <- "?published_date=2018-01-01"

nyt.url <- str_c(host, path, query, api_key, sep = "", collapse = NULL)

raw.result2 <- GET(nyt.url, verbose())
```


```{r}
http_status(raw.result2)
```

## Creating the Bestsellers Dataframe

Now, to create a dataframe of the Bestselling Books on Jan 1, 2018. 

First, extracting the content from the body as a JSON, then reading it into R before furthering munging. 

```{r}
this.raw.content2 <- content(raw.result2, "text", encoding='UTF-8')

this.json2 <- fromJSON(this.raw.content2, flatten = TRUE)

bestsellers <- this.json2$results$list
```

Here we are only concerned with the data in the books column. 

```{r}
names(bestsellers[[1]])
```


### Method 1, Works in RStudio, but doesn't when Knitting

I then follow this procedure:

(1) Unlisting the list of books into a vector;
(2) From the list of lists books, forming a matrix from the transposed vector with rows equal to the number the books in the list, and columns equal to the number of book attributes.
(3) Converting this matrix into a dataframe with column headings of the four attributes that I want to draw out.  


```{r}
books <- unlist(bestsellers[[1]]["books"])
books <- matrix(t(books), nrow = 29, ncol = 5)
books_df <- data.frame("Title" = books[21,], "Author" = books[4,], "Description" = books[12,], "Wks_Bestseller" = books[23,])
books_df

```

### Method 2, Doesn't work in RStudio, but does when Knitting

Note, these next lines of code in the jsonlite vignette did not work for me in RMarkdown due to: Error in bestsellers[[1, "books"]] : incorrect number of subscripts. See sources below. But, after knitting, this code mysterioudly works, though it didn't in RStudio. 

```{r}
category1 <- bestsellers[[1, "books"]]
df <- subset(category1, select = c("title", "author", "description", "weeks_on_list"))
pander(df, style = 'rmarkdown')
```



The following links were helpful to this assignment.

Sources:

https://www.programmableweb.com/news/how-to-access-any-restful-api-using-r-language/how-to/2017/07/21?page=2

https://www.r-bloggers.com/accessing-apis-from-r-and-a-little-r-programming/

https://rstudio-pubs-static.s3.amazonaws.com/223073_37195ccc31b846fc8c963e5d10416887.html

https://cran.r-project.org/web/packages/jsonlite/vignettes/json-apis.html

