---
title: "CUNY DATA607_Wk7_Zherold_Scraping"
output: html_document
---

In this assignment, I load Notepad-generated HTML, XML, and JSON files, which were uploaded to Github, into R as dataframes. 

```{r, message=FALSE,warning=FALSE}
library(XML)
library(xml2)
library(jsonlite)
library(httr)
library(knitr)
```

#### HTML Table >> R, using httr's GET ()... 


```{r}
dfhtml <- GET("https://raw.githubusercontent.com/ZacharyHerold/CUNY-DATA607/master/books_html.html")
dfhtml 
```


```{r}
class(dfhtml)
```


#### ... then converting it into a dataframe with readHTMLTable()

Not entirely sure why I had to repeat the GET() line of code again here in this chunk. When I don't have it， I get Error in rawToChar(dfhtml$content) : argument 'x' must be a raw vector

```{r}
## Load HTML Table
dfhtml <- GET("https://raw.githubusercontent.com/ZacharyHerold/CUNY-DATA607/master/books_html.html")
dfhtml <- as.data.frame(readHTMLTable(rawToChar(dfhtml$content), stringsAsFactors = F))
names(dfhtml) <- c("Title", "Author_First", "Author_Last", "DoP")
kable(dfhtml) 
```

```{r}
class(dfhtml)
```
 
 Further manipulations are required to separate the multiple authors. One possible way could involve, first, creating NULL Author_First2 and Author_Last2 columns, then using stringr to do strsplit the Author columns at the comma, after they had been unlisted.  


#### XML Document >> R, using GET() and XMLParse()...

* The httr package did prove more reliable than RCurl when Knitting the Markdown file to html. The readHTMLTable was missing function calls (Handlers, etc.) when using getURL().


```{r}
## Load XML File
url.xml <- GET("https://raw.githubusercontent.com/ZacharyHerold/CUNY-DATA607/master/books_xml.xml")
dfxml <- xmlParse(url.xml)
dfxml
```


#### ...and then xmlToDataFrame()

```{r}
books.df <- xmlToDataFrame(dfxml, stringsAsFactors = FALSE)
kable(books.df)
```

Should debug XML file as follows:

    <author>
      <first_name>Steve</first_name>
      <last_name>Fraser</last_name>
    </author>  
    <author>
      <first_name>Gary</first_name>
      <last_name>Gerstle</last_name>
    </author>
    


#### JSON Document >> R, using httr's GET()

```{r}
## Load JSON File
url.json <- GET("https://raw.githubusercontent.com/ZacharyHerold/CUNY-DATA607/master/books_json.json")
url.json
```

#### .... readLines(), editing out the garbage at the beginning of the JSON string...
.
```{r}
url.json <- "https://raw.githubusercontent.com/ZacharyHerold/CUNY-DATA607/master/books_json.json"
json.data <- readLines(url.json)
json.data[[1]] <- "{\"social economy books\" :["
```

Checking the class of the readLine's output. Had difficulty without readLines using the fromJSON() function. 

```{r}
class(json.data)
```

#### .... and finally coercing to a dataframe with the as.data.frame() function. 

```{r results = 'asis'}
dfjson <- fromJSON(json.data)
dfjson <- as.data.frame(dfjson)
names(dfjson) <- c("Title","Author_First","Author_Last","Year","Pages","Publisher")
kable(dfjson)
```

